---
title: "ggplot2-STAA566"
author: "Nat Koonmee"
date: '2022-09-21'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = F)

library("readxl")
library(dplyr)
library(ggpubr)
library(tidyverse)
library(ggplot2)
library(gapminder)
library(TSstudio)
library(hrbrthemes)
library(scales)
library(lubridate)
library(caret)
```


## Header

This is Fort Collins temperature report datasets from https://ccc.atmos.colostate.edu/~autowx/fclwx_access.php \newline
We could see from the table below that in each day, there are 24 temperature reports which mean inorder to get a daily temperature, we need to conbind those hourly to daily.\newline
The last columns on the datasets is a West Nile Virus report from https://cdphe.colorado.gov/animal-related-diseases/west-nile-virus/west-nile-virus-and-your-health which is also, an open source.

```{r}
df = read.csv("DF_allcases.csv")
df <- df %>% select(2,3,4,6,7,9,11,12,13,15,16,17,18,19,20)
name <- c("index", "date", "time", "TempC", "Humidity", "DewPointC", "WindMS", "WindDir", "GustMS", "GustDir", "Pressure", "Solar", "PrecIN", "PrecMM", "WNVCasesReport")
colnames(df) <- name
df$date <- as.Date(df$date, format = "%m/%d/%Y")
head(df)

```

```{r}
# colors <- c("Avg Temp." = "cyan4", "Min Temp." = "cornflowerblue", "Max Temp." = "brown1")
# 
# dftemp <- df %>% select(c(2,4))
# tempavg <- aggregate(dftemp[2], by= dftemp[1], mean)
# tempmin <- aggregate(dftemp[2], by= dftemp[1], min)
# tempmax <- aggregate(dftemp[2], by= dftemp[1], max)
# tempdata <- data.frame(Date_MST = tempavg[,1], 
#                        Average_temperature = tempavg[,2],
#                        Min_temperature = tempmin[,2],
#                        Max_temperature = tempmax[,2])
# 
# p <- ggplot(tempdata) + 
#   # geom_line(aes(x = Date_MST, y = Average_temperature), color = "Avg Temp.") + 
#   # geom_line(aes(x = Date_MST, y = Min_temperature), color = "Min Temp.") +
#   # geom_line(aes(x = Date_MST, y = Max_temperature), color = "Max Temp.") + theme_bw() + 
#   geom_line(aes(x = Date_MST, y = Average_temperature), color = "cornflowerblue") + 
#   geom_line(aes(x = Date_MST, y = Min_temperature), color = "cyan4") +
#   geom_line(aes(x = Date_MST, y = Max_temperature), color = "brown1") + 
#   labs(x = "year", y = "Temperature", color = "Legend") + scale_color_manual(values = colors) 
# 
# 
# p
```


```{r}
colors <- c("AvgTemp" = "cyan4", "MinTemp" = "cornflowerblue", "MaxTemp" = "brown1")

dftemp <- df %>% select(c(2,4))
tempavg <- aggregate(dftemp[2], by= dftemp[1], mean)
tempmin <- aggregate(dftemp[2], by= dftemp[1], min)
tempmax <- aggregate(dftemp[2], by= dftemp[1], max)
tempdata <- data.frame(Date_MST = tempavg[,1], 
                       Average_temperature = tempavg[,2],
                       Min_temperature = tempmin[,2],
                       Max_temperature = tempmax[,2])

p <- ggplot(tempdata, aes(x = Date_MST)) + 
  geom_line(aes(y = Average_temperature), colour  = 'cyan4') +
  geom_line(aes(y = Min_temperature), colour   = "cornflowerblue") +
  geom_line(aes(y = Max_temperature), colour  = 'brown1')  +
  labs(x = "year", y = "Temperature (C)", color = "Legend", title = "Daily Temperature at Fort Collins from 2002 to 2022") +
  scale_colour_manual(name = "ll",
    values = c("cyan4" = "cyan4", "cornflowerblue" = "cornflowerblue", "brown1" = "brown1"),
                      labels = c("l1", "l2", "l3")) +
  theme_bw() 


p


```

We could see from this plot, that looking for daily temp. plot does not help anything much, since there are too many information and noise.


```{r}
# yearly <- tempdata %>%
#   mutate(date = year(Date_MST)) %>%
#   group_by(date) %>%
#   summarize(average = mean(Average_temperature), min = min(Average_temperature), max = max(Average_temperature))
# p <- ggplot(yearly) + 
#   geom_line(aes(x = date, y = average), color = "cyan4") + 
#   geom_line(aes(x = date, y = min), color = "cornflowerblue") +
#   geom_line(aes(x = date, y = max), color = "brown1") + theme_bw()
# 
# 
# p

```


```{r}
yearly <- tempdata %>%
  mutate(date = year(Date_MST)) %>%
  group_by(date) %>%
  summarize(average = mean(Average_temperature), min = mean(Min_temperature), max = mean(Max_temperature))
p <- ggplot(yearly) + 
  geom_line(aes(x = date, y = average), color = "cyan4") + 
  geom_line(aes(x = date, y = min), color = "cornflowerblue") +
  geom_line(aes(x = date, y = max), color = "brown1") + theme_bw() +
  labs(x = "year", y = "Temperature (C)", title = "Yearly Temperature at Fort Collins from 2002 to 2022")


p
```

Instead of plotting daily temperature, we still do not receive a lot of information from yearly Temperature, it is still almost a flat line, and there is no pattern or any correlation.


```{r}
name_month = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", 'Aug', 'Seb', "Oct", "Nov", "Dec")
 month <- tempdata %>%
  mutate(date = month(Date_MST)) %>%
  group_by(date) %>%
  summarize(average = mean(Average_temperature), min = min(Average_temperature), max = max(Average_temperature))
p <- ggplot(month) + 
  geom_line(aes(x = date, y = average), color = "cyan4") + 
  geom_line(aes(x = date, y = min), color = "cornflowerblue") +
  geom_line(aes(x = date, y = max), color = "brown1") + theme_bw()+ 
  labs(x = "Month", y = "Temperature(C)")
  # scale_x_discrete(labels = name_month)

p

p2 <- ggplot(month, aes(x = date, y = average)) + 
  geom_point() + geom_line() +
  geom_smooth(method = "loess") + theme_bw()

p2


```

Now, we can see a pattern of average Temp by month through 20 years, we could see that during month 6 - 7 or June - July is the peak of the Temp. in Fort Collins.

```{r}


tempdata["Year"] = ''
tempdata["Month"] = ""
for (i in 1:length(tempdata[,1])){
  tempOutput = format(as.Date(tempdata[i, 1], format="%Y/%m/%d"),"%Y")
  tempdata[i, 5] = tempOutput
  
  tempdata[i, 6] = format(as.Date(tempdata[i, 1], format="%Y/%m/%d"),"%m")
}



 # Yearly <- tempdata %>%
 #  group_by(Year) %>%
 #  summarize(average = mean(Average_temperature), min = mean(Min_temperature), max = mean(Max_temperature))
 

 
 p <- ggplot(tempdata) + 
  geom_boxplot(aes(x = Month, y = Average_temperature)) + theme_bw()

p


```

```{r}

```


```{r}
ggplot(df, aes(x = Humidity)) + geom_histogram(aes(y = ..density..),binwidth = 3, color ="black", fill = "white") + geom_density(alpha=.2, fill="#FF6666") +
  theme_bw() +
  labs(title = "Histogram or Humidity from 2002 to 2020")
  
df1 <- df
df1["Humidity > 50"] = "No"
df1["Year"] = ''
df1["Month"] = ""

for (i in 1:length(df1[,1])){
  # temp = "No"
  if (df1[i,5] < 50){
    df1[i,16] = "Yes"
  }
  
  tempOutput = format(as.Date(df1[i, 2], format="%Y/%m/%d"),"%Y")
  df1[i, 17] = tempOutput
  
  df1[i, 18] = format(as.Date(df1[i, 2], format="%Y/%m/%d"),"%m")
  
}  
  

```







```{r}
weather2022 <- df1 %>% filter(Year == "2022")


ggplot(weather2022, aes(x = weather2022$GustMS, y = weather2022$WindMS, color = weather2022$`Humidity > 50`, fill = weather2022$`Humidity > 50`)) + geom_point(alpha = .3) + labs(x = "Guest Speed Meter per second", y = "Wind Speed Meter per second", colour = "is Humidty greater than 50")

```

```{r}
weather2022 <- df1 %>% filter(Year == "2022" & WindMS < 3 & GustMS < 7.5)


ggplot(weather2022, aes(x = weather2022$GustMS, y = weather2022$WindMS, color = weather2022$`Humidity > 50`, fill = weather2022$`Humidity > 50`)) + geom_point(alpha = .3) + labs(x = "Guest Speed Meter per second", y = "Wind Speed Meter per second", colour = "is Humidty greater than 50")
```





















































